---
title: 'Assignment #3 - PALMA'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)         # for graphing and data cleaning
library(tidymodels)        # for modeling
library(themis)            # for step functions for unbalanced data
library(doParallel)        # for parallel processing
library(stacks)            # for stacking models
library(naniar)            # for examining missing values (NAs)
library(lubridate)         # for date manipulation
library(moderndive)        # for King County housing data
library(vip)               # for variable importance plots
library(patchwork)         # for combining plots nicely
library(ranger)
library(xgboost)
theme_set(theme_minimal()) # Lisa's favorite theme
```

```{r data}
data("lending_club")
# Data dictionary (as close as I could find): https://www.kaggle.com/wordsforthewise/lending-club/discussion/170691
```

## Put it on GitHub!        

**Task**: When you are finished with the assignment, post a link below to the GitHub repo for the assignment. Make sure the link goes to a spot in the repo where I can easily find this assignment. For example, if you have a website with a blog and post the assignment as a blog post, link to the post's folder in the repo. As an example, I've linked to my GitHub stacking material [here](https://github.com/llendway/ads_website/tree/master/_posts/2021-03-22-stacking).

Website Link: **https://github.com/apalma127/assignment-3**

## Modeling

The outcome we are interested in predicting is `Class`. And according to the dataset's help page, its values are "either 'good' (meaning that the loan was fully paid back or currently on-time) or 'bad' (charged off, defaulted, or 21-120 days late)".

**Tasks:** 

1. Explore the data, concentrating on examining distributions of variables and examining missing values. 

**Quantitative Variables**

```{r}
lending_club %>% 
  select(where(is.numeric)) %>% 
  pivot_longer(cols = everything(),
               names_to = "variable", 
               values_to = "value") %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(vars(variable), 
             scales = "free")
```

Points of Interest:

-
-
-


**Categorical Variables**

```{r}
lending_club %>%
select(where(is.factor)) %>% 
  pivot_longer(cols = everything(),
               names_to = "variable", 
               values_to = "value") %>% 
  ggplot(aes(x = value)) +
  geom_bar() +
  facet_wrap(vars(variable), 
             scales = "free", 
             nrow = 2)
```

Points of Interest:

-
-
-

```{r}
lending_club %>%
  count(sub_grade)
```


2. Split the data into training and test, putting 75\% in the training data. Stratify by `Class` (add `strata = `Class` to the `initial_split()` function).

```{r}
set.seed(494) # for reproducibility

lending_split <- initial_split(lending_club, 
                             prop = .75, starta = "Class")

lending_training <- training(lending_split)
lending_test <- testing(lending_split)
```


3. Set up the recipe and the pre-processing steps to build a lasso model. Some steps you should take:

* Use `step_upsample()` from the `themis` library to upsample the "bad" category so that it is 50\% of the "good" category. Do this by setting `over_ratio = .5`.

* Use `step_downsample()` from the `themis` library to downsample the "good" category so the bads and goods are even - set `under_ratio = 1`. Make sure to do this step AFTER `step_upsample()`.  

* Make all integer variables numeric (I'd highly recommend using `step_mutate_at()` and using the `all_numeric()` helper or this will be a lot of code). This step might seem really weird right now, but we'll want to do this for the model interpretation we'll do in a later assignment.  

* Think about grouping factor variables with many levels.  

Once you have that, use `prep()`, `juice()`, and `count()` to count the number of observations in each class. They should be equal. This dataset will be used in building the model, but the data without up and down sampling will be used in evaluation.

```{r}
set.seed(456)

lasso_recipe <- recipe(Class ~ ., 
                       data = lending_training) %>% 
  step_upsample(Class, over_ratio = 0.5) %>%
  step_downsample(Class, under_ratio = 1) %>%
  step_mutate_at(all_numeric(), 
                fn = ~ as.numeric(.)) %>%
  step_mutate(sub_grade = as.character(sub_grade),
                sub_grade =
                        case_when(
                          sub_grade %in% paste("A",1:6, sep = "") ~ "A",
                          sub_grade %in% paste("B",1:6, sep = "") ~ "B",
                          sub_grade %in% paste("C",1:6, sep = "") ~ "C",
                          sub_grade %in% paste("D",1:6, sep = "") ~ "D",
                          sub_grade %in% paste("E",1:6, sep = "") ~ "E",
                          sub_grade %in% paste("F",1:6, sep = "") ~ "F",
                          sub_grade %in% paste("G",1:6, sep = "") ~ "G",
                          TRUE ~ sub_grade), 
              sub_grade = as.factor(sub_grade)) %>%
  step_dummy(all_nominal(), 
             -all_outcomes()) %>% 
  step_normalize(all_predictors(), 
                 -all_nominal())
```

```{r}
lasso_recipe %>% 
  prep(lending_training) %>%
  juice() 
```


4. Set up the lasso model and workflow. We will tune the `penalty` parameter.

```{r}
lasso_mod  <- 
  logistic_reg(mixture = 1) %>% 
  set_engine("glmnet") %>% 
  set_args(penalty = tune()) %>% 
  set_mode("classification")

lasso_wf <-  workflow() %>% 
  add_recipe(lasso_recipe) %>% 
  add_model(lasso_mod)
```


5. Set up the model tuning for the `penalty` parameter. Be sure to add the `control_stack_grid()` for the `control` argument so we can use these results later when we stack. Find the accuracy and area under the roc curve for the model with the best tuning parameter.  Use 5-fold cv.

Tune:

```{r}
set.seed(494) #for reproducible 5-fold
lending_cv <- vfold_cv(lending_training, v = 5)


penalty_grid <- grid_regular(penalty(),
                             levels = 10)

ctrl_grid <- control_stack_grid()

lasso_tune <-  
  lasso_wf %>% 
  tune_grid(
    resamples = lending_cv,
    grid = penalty_grid,
    control = ctrl_grid)

lasso_tune %>% 
  select(id, .metrics) %>% 
  unnest(.metrics) %>% 
  filter(.metric == "accuracy")

```


```{r}

lasso_tune %>% 
  collect_metrics() %>% 
  filter(.metric == "accuracy") %>% 
  ggplot(aes(x = penalty, y = mean)) +
  geom_point() +
  geom_line() +
  scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10",scales::math_format(10^.x))) +
  labs(x = "penalty", y = "accuracy")
```

```{r}
lasso_tune %>% 
  show_best(metric = "accuracy")
```

```{r}
best_param <- lasso_tune %>% 
  select_best(metric = "accuracy")
best_param
```

```{r}
lasso_tune %>%
  select_best(metric = "accuracy")

lasso_tune %>%
  show_best(metric = "accuracy") %>%
  filter(.config == "Preprocessor1_Model09")

lasso_tune %>%
  show_best(metric = "roc_auc") %>%
  filter(.config == "Preprocessor1_Model09")
```
**ACCURACY = 0.7336347**

**ROC AUC = 0.7487696 **


6. Set up the recipe and the pre-processing steps to build a random forest model. You shouldn't have to do as many steps. The only steps you should need to do are making all integers numeric and the up and down sampling. 

```{r}
set.seed(456)
rf_recipe <-  
  recipe(Class ~ ., 
        data = lending_training) %>% 
  step_upsample(Class, over_ratio = 0.5) %>%
  step_downsample(Class, under_ratio = 1) %>%
  step_mutate_at(all_numeric(), 
                fn = ~ as.numeric(.))
  
```


7. Set up the random forest model and workflow. We will tune the `mtry` and `min_n` parameters and set the number of trees, `trees`, to 100 (otherwise the next steps take too long).

```{r}
rf_model <- rand_forest(mtry = tune(), 
              min_n = tune(), 
              trees = 100) %>% 
  set_mode("classification") %>% 
  set_engine("ranger")

rf_workflow <- workflow() %>% 
  add_recipe(rf_recipe) %>% 
  add_model(rf_model) 
```

8. Set up the model tuning for both the `mtry` and `min_n` parameters. Be sure to add the `control_stack_grid()` for the `control` argument so we can use these results later when we stack. Use only 3 levels in the grid. For the `mtry` parameter, you need to put `finalize(mtry(), lending_training %>% select(-Class))` in as an argument instead of just `mtry()`, where `lending_training` is the name of your training data. This is because the `mtry()` grid will otherwise have unknowns in it. This part can take a while to run.

```{r}
rf_penalty_grid <- grid_regular(
  finalize(mtry(), lending_training %>% select(-Class)),
  min_n(),
  levels = 3)


rf_tune <- 
  rf_workflow %>% 
  tune_grid(
    resamples = lending_cv, 
    grid = rf_penalty_grid, 
    control = control_stack_grid())

```

9. Find the best tuning parameters. What are the accuracy and area under the ROC curve for the model with those tuning parameters?

```{r}

rf_tune %>%
  select_best(metric = "accuracy")

rf_tune %>%
  show_best(metric = "accuracy") %>%
  filter(.config == "Preprocessor1_Model2")

rf_tune %>%
  show_best(metric = "roc_auc") %>%
  filter(.config == "Preprocessor1_Model2")

```


**Accuracy = 0.924379**

**ROC AUC = 0.7257118**


10. Next, we will fit a boosted tree using xgboost. We will only tune the `learn_rate` parameter. I have specified the model, recipe, and workflow below already (uncomment the code - you can this by highlighting it and then in the code tab at the top, choose comment/uncomment lines). You need to set up a grid of ten values for the tuning parameter and tune the model. Be sure to add the `control_stack_grid()` for the `control` argument so we can use these results later when we stack.

```{r}
xgboost_spec <-
  boost_tree(
    trees = 1000,
    min_n = 5,
    tree_depth = 2,
    learn_rate = tune(),
    loss_reduction = 10^-5,
    sample_size = 1) %>%
  set_mode("classification") %>%
  set_engine("xgboost")

xgboost_recipe <- recipe(formula = Class ~ ., data = lending_training) %>%
  step_upsample(Class, over_ratio = .5) %>%
  step_downsample(Class, under_ratio = 1) %>%
  step_mutate_at(all_numeric(),
                 fn = ~as.numeric(.)) %>%
  step_novel(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_zv(all_predictors())

xgboost_workflow <-
  workflow() %>%
  add_recipe(xgboost_recipe) %>%
  add_model(xgboost_spec)

set.seed(494)
registerDoParallel() 

boost_penalty_grid <- grid_regular(
  learn_rate(),
  levels = 10)

boost_tune <- xgboost_workflow %>% 
    tune_grid(
    resamples = lending_cv, 
    grid = boost_penalty_grid, 
    control = control_stack_grid())
```

11. Find the best tuning parameters. What are the accuracy and area under the ROC curve for the model with those tuning parameters?

```{r}
boost_tune %>%
  select_best(metric = "accuracy")

boost_tune %>%
  collect_metrics(metric = "roc_auc") %>%
  filter(.config == "Preprocessor1_Model10")

```

**ACCURACY = 0.8051965**

**ROC AUC = 0.6943527**

12. Create a model stack with the candidate models from the previous parts of the exercise and use the `blend_predictions()` function to find the coefficients of the stacked model. Create a plot examining the performance metrics for the different penalty parameters to assure you have captured the best one. If not, adjust the penalty. (HINT: use the `autoplot()` function). Which models are contributing most?

```{r}
lending_stack <- stacks() %>% 
  add_candidates(lasso_tune) %>% 
  add_candidates(rf_tune) %>% 
  add_candidates(boost_tune)
```

```{r}
lending_blend <- lending_stack %>%
            blend_predictions()

lending_blend
```


**The greatest contributors to the stacking are the log_reg lasso tune by the most followed by random forest tune (whose impact is half of that of the log reg lasso).**


```{r}
lending_blend$metrics %>% 
  filter(.metric == "accuracy")
```

```{r}
autoplot(lending_blend)

```

**No adjustment necessary as the num of members is minimized all while maximizing accuracy and roc auc**


13. Fit the final stacked model using `fit_members()`. Apply the model to the training data. Compute the accuracy, construct a confusion matrix, and create a density plot with `.pred_good` on the x-axis (the probability of a response of "good"), filled by `Class`. Comment on what you see. 

```{r}
lending_final_stack <- lending_blend %>%
                    fit_members(data = lending_training)
```


14. In the previous problem, you saw that although the accuracy was quite high, the true negative rate (aka sensitivity) was terrible. It's common to see this when one of the classes has low representation. What we want to do now is investigate what happens in each of our models. Below I've provided code to investigate the lasso model (where `lasso_tune` is the name of my tuning step). Do similar things for the random forest and xgboost models. If you'd like to have a better true negative rate, which models would you choose and how would you go about doing this in a less manual way (you don't need to write code to do it - just describe it in words). Be sure to remove the `eval=FALSE` when you are finished.

```{r, eval=FALSE}
lasso_tune %>% 
  collect_predictions() %>% 
  group_by(id, penalty) %>% 
  summarize(accuracy = sum((Class == .pred_class))/n(),
            true_neg_rate = sum(Class == "bad" & .pred_class == "bad")/sum(Class == "bad"),
            true_pos_rate = sum(Class == "good" & .pred_class == "good")/sum(Class == "good")) %>% 
  group_by(penalty) %>% 
  summarize(across(accuracy:true_pos_rate, mean))
```


## Shiny app

Shiny links on our course [Resource](https://advanced-ds-in-r.netlify.app/resources.html)

**Tasks:**

In the future, you are going to create an app that allows a user to explore how the predicted probability of a loan being paid back (or maybe just the predicted class - either "good" or "bad") changes depending on the values of the predictor variables.

For this week, I want you to answer the following questions:

1. How can you save a model you built to use it later (like in the shiny app you'll create)?

**Just like from shiny using anim_save(), with a model you can use saveRDS (I think?) or just save -->  save(model, file="model.Rdata") ... which can be brought back with the load function**

2. For shiny apps that get published (like yours will), it's very important to have ALL the libraries that are used within the app loaded. If we were going to use the stacked model, which libraries do you think we'd need to load in our app?  

**I would imagine we need tidyverse, tidymodels, DEFINITELY stacks, depending on the model --glmnet, ranger, kknn; naniar, lubridate, vip**


3. You'll want the user to be able to choose values for each variable in the model. How will you come up with the values they can choose for quantitative and categorical data? Give one example for each, either using code or in words.  


**I can use graphical processes like the preprocessing work we have done for quantitative and categorical variables. By doing so, I cna eye ball right off the bat some ranges of inputs for things, especially binary outputs.  I cna use the count function for categorical variables to see the breakdown and fro quantitative I can sort the data and/or simply find the min and max values from the category to establish a user input range.**


4. You will need to populate each variable with an initial value. Which value will you choose? Is there a nice way to do this programatically (ie. with code)?


**For each variable, I think it would make sense to start with the neutral case and go from there to avoid any appearance of guiding the user.  This would mean all categories that are extra and that come after receiving a loan would be set to 0 ie delinq variables, total_bal, open_il...others like funded amt, interest rate, and annual income will be randomly generated in a reasonable range allowing for the other variables to play off that.  Programmatically, we can set some to 0 at the start easily and others can simply be randomized using random generator**


## Coded Bias

We will be watching some of the [Coded Bias](https://www.codedbias.com/) film together on Thursday. It is streaming on Netflix. Write a short reflection. If you want some prompts, reflect on: 

Was there a part that surprised you and why? 

**_________________________________**


**REMEMBER TO ADD YOUR GITHUB LINK AT THE TOP OF THE PAGE AND UNCOMMENT THE `knitr` OPTIONS.**


